### 2008-12-21 talk/ecoop2009/conclusion.tex
some minore changes
add conclusion and future work

\section{Conclusion}

XXX add me

>>>>

\section{Conclusion and Future Work}

In this paper we presented PyPy's JIT compiler generator, based on partial
evaluation techniques, which can automatically turn an interpreter into a JIT
compiler, requiring the language developers to only add few \texttt{hint}s to
guide the generation process.

We showed that classical partial evaluation cannot remove all the overhead
proper of dynamically typed languages, and how the new operation called
\emph{promotion} solves the problem, by delaying compile-time until the JIT
knows enough to produce efficient code, and by continuously intermixing
compile-time and runtime.  Moreover, we showed that our simple but still
practically useful technique to avoid allocation of intermediate unnecessary
objects plays well with promotion and helps to produce even better code.

Finally, we presented the CLI backend for PyPy's JIT compiler generator, whose
goal is to produce .NET bytecode at runtime.  We showed how it is possible to
circumvent intrinsic limitations of the virtual machine to implement
promotion.  As a result, we proved that the idea of \emph{JIT layering} is
worth of further exploration, as it makes possible for dynamically typed
languages to be even faster than their statically typed counterpart in some
circumstances.

As a future work, we want to explore different strategies to make the frontend
producing less code, maintaining comparable or better performances.  In
particular, we are working on a way to automatically detect loops in the user
code, as tracing JITs do \cite{gal_hotpathvm_2006}.  By compilining whole
loops at once, the backends should be able to produce better code than today.

At the moment, some bugs and minor missing features prevent the CLI JIT
backend to handle more complex languages such as Python and Smalltalk.  We are
confident that once these problems will be fixed, we will get performance
results comparable to TLC, as the other backends already demonstrate
\cite{PyPyJIT}.  Moreover, if the current implementation of flexswitches will
prove to be too slow for some purposes, we want to explore alternative
implementation strategies, also considering the new features that might be
integrated into virtual machines.
